---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(scales)
library(here)
```


# Read in and explore the data
```{r}
day_visit_location <- read_csv(here("raw_data/tourism_day_visits_location.csv")) %>% 
  clean_names()
location_data <- read_csv(here("raw_data/Datazone2011lookup.csv")) %>% 
  clean_names()
regional_domestic_tourism <- read_csv(here("raw_data/regional_domestic_tourism.csv")) %>% 
  clean_names()
day_visit_activities <- read_csv(here("raw_data/tourism_day_visits_activities.csv")) %>% 
  clean_names()
day_visit_demographics <- read_csv(here("raw_data/tourism_day_visits_demographics.csv")) %>% 
  clean_names()
day_visit_transport <- read_csv(here("raw_data/tourism_day_visits_transport.csv")) %>% 
  clean_names()
accomm_occupancy <- read_csv(here("raw_data/scottish_accomodation_occupancy.csv")) %>% 
  clean_names()
```


# Activities: What kind of tourism activity generates the highest income for Scotland? Which kind of activity generates the most visits to sites?

## Plot the hightest ranking activities

```{r}
# create names for the labels
activity_names <- c("Meal Out", "Shopping", "Visit Family or Friends", "Night Out", "Outdoor Activity", "General day out")

# select the top activities
day_visit_activities %>% 
  filter(tourism_activity %in% c("Went out for a meal", "Visited friends or family for leisure", "Shopping for items that you do not regularly buy", "Night out to a bar, pub and/or club", "Outdoor leisure activities e.g. walking, golf", "General day out/ to explore an area")) %>% 
  group_by(tourism_activity, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(tourism_activity, -sum), y = sum, fill = units), position = "dodge") +
  scale_x_discrete(labels= activity_names) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Activity",
       y = "Millions",
       fill = "",
       title = "Top Tourist Activities",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```
 ## Plot the highest ranking activities by visits only
```{r}
activity_names_visits <- c("Visit Family or Friends", "Meal Out", "Outdoor Activity", "General day out", "Night Out", "Shopping")

day_visit_activities %>% 
  filter(tourism_activity %in% c("Went out for a meal", "Visited friends or family for leisure", "Shopping for items that you do not regularly buy", "Night out to a bar, pub and/or club", "Outdoor leisure activities e.g. walking, golf", "General day out/ to explore an area"),
         units == "Million Visits") %>% 
  group_by(tourism_activity, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(tourism_activity, -sum), y = sum), fill = "#B27300") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", size = 10),
        title =element_text(size=12, face='bold'),
        axis.title=element_text(size=12)) +
  labs(x = "Activity",
       y = "No. of Visits (m)\n",
       title = "Top Tourist Activities by Visits",
       subtitle = "2013 - 2019") +
  scale_x_discrete(labels= activity_names_visits)
  
```



# Demographics
## create visualisations for each category in the demographis data
```{r}
 day_visit_demographics %>% 
  filter(age != "All") %>% 
  group_by(age, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(age, -sum), y = sum, fill = units), position = "dodge") +
   theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Age Group",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Age",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

```{r}
day_visit_demographics %>% 
  filter(marital_status != "All")%>% 
  group_by(marital_status, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(marital_status, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Marital Staus",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Marital Status",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

```{r}
day_visit_demographics %>% 
  filter(gender != "All") %>% 
  group_by(gender, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(gender, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Gender",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Gender",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

```{r}
# create list of employment status 
employed_status <- c("Employed/Self-employed", "Unemployed", "Full/Part time education")

day_visit_demographics %>% 
  filter(employment_status != "All") %>% 
  group_by(employment_status, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(employment_status, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Employment Status",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Employment Status",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300")) +
  scale_x_discrete(labels= employed_status)
```

```{r}
day_visit_demographics %>% 
  filter(children != "All") %>% 
  group_by(children, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(children, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Family Status",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Family Status",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

```{r}
day_visit_demographics %>% 
  filter(access_to_car != "All") %>% 
  group_by(access_to_car, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(access_to_car, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Access to Car",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Car Accessibility",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

```{r}
day_visit_demographics %>% 
  filter(social_grade != "All") %>% 
  group_by(social_grade, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(social_grade, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Social Grade",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Social Grade",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```
## change the dataframe to a longer version so that totals for each subcategory can be calculated
```{r}
demographics_long <- day_visit_demographics %>% 
  pivot_longer(cols = c(age:social_grade), names_to = "category", values_to = "status") %>% 
  filter(status != "All") %>% 
  arrange(date_code, breakdown_of_domestic_tourism, category)
```

# calculate totals of expenditure for each subcategory
```{r}
demograhics_exp_total <- demographics_long %>% 
  group_by(status, breakdown_of_domestic_tourism) %>% 
  filter(breakdown_of_domestic_tourism == "Expenditure") %>% 
  summarise(sum = sum(value)) %>% 
  arrange(desc(sum))
```



# Locations of day visits
## create a summary of exp/visits by location
```{r}
location_summary <- day_visit_location %>% 
  group_by(type_of_location_visited,units) %>% 
  summarise(sum = sum(value)) %>% 
  pivot_wider(names_from = units, values_from = sum) %>% 
  clean_names() %>% 
  arrange(desc(million_pounds_gbp))

location_summary
  
```
## Plot the expenditure and visits by location type
```{r}
day_visit_location %>% 
  filter(type_of_location_visited != "All areas") %>% 
  group_by(type_of_location_visited,units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(type_of_location_visited, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 25, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Location Type",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Location Type",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```


# Transport

## create a summary of exp/visits by transport
```{r}
day_visit_transport %>% 
  group_by(method_of_travel, units) %>% 
  summarise(sum = sum(value)) %>% 
  pivot_wider(names_from = units, values_from = sum) %>% 
  clean_names() %>% 
  arrange(desc(million_pounds_gbp))
```


## Plot exp/visits by transport
```{r}
day_visit_transport %>% 
  filter(method_of_travel != "All") %>% 
  group_by(method_of_travel,units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(method_of_travel, -sum), y = sum, fill = units), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Mode of Transport",
       y = "Millions",
       fill = "",
       title = "Expenditure and Visits by Mode of Transport",
       subtitle = "2013 - 2019") +
  scale_fill_manual(labels = c("Pounds (GBP)", "Visits"), values = c("million pounds (GBP)" = "#003300", "Million Visits" = "#B27300"))
```

# Regional Tourism
```{r}
# look for the distinct location codes
glimpse(regional_domestic_tourism)

regional_domestic_tourism %>% 
  distinct(feature_code)
```

# summarise the data, taking out the feature_code for all of Scotland
```{r}
regional_summary <- regional_domestic_tourism %>% 
  filter(feature_code != "S92000003") %>% 
  group_by(region_of_residence ,units) %>% 
  summarise(sum = sum(value))

```


```{r}
# extract local area codes from location code data and rename to match tourism data 
location_code <- location_data %>% 
  rename(feature_code = la_code,
         area_visited = la_name) %>% 
  select(c(feature_code, area_visited))
```


```{r}
# add local area name to the tourism data and remove the data for all of Scotland
regional_visits <- location_code %>% 
  filter(feature_code != "S92000003") %>% 
  distinct() %>% 
  right_join(regional_domestic_tourism, by = "feature_code")
```


## create graph for total expenditure by region of residence
```{r}
regional_visits %>% 
  filter(units == "million pounds (GBP)") %>% 
  ggplot() +
  geom_col(aes(x = reorder(region_of_residence, -value), y = value), fill = "#003300", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Region of Residence",
       y = "Millions",
       fill = "",
       title = "Expenditure by region of tourist's residence",
       subtitle = "2013 - 2019")
```
## create graph for total expenditure by region of residence and area visited
```{r}
regional_visits %>% 
  drop_na(area_visited) %>% 
  group_by(area_visited) %>% 
  filter(region_of_residence != "All of GB",
         units == "million pounds (GBP)") %>%
  ggplot() +
  geom_col(aes(x = reorder(area_visited, -value), y = value, fill = region_of_residence), position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 55, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Area Visited",
       y = "Millions",
       fill = "",
       title = "Expenditure in Regional Areas by Tourist's Residence",
       subtitle = "2013 - 2019") +
  scale_fill_manual(values = c("England" = "#B20000", "Scotland" = "#211551"))
```
## create graph for total nights stayed and visits by region of residence for top areas visited
```{r}

regional_visits %>% 
  drop_na(area_visited) %>% 
  group_by(area_visited) %>% 
  filter(region_of_residence != "All of GB",
         area_visited %in% c("City of Edinburgh", "Glasgow City", "Dumfries and Galloway", "Highland", "Argyle and Bute", "Perth and Kinross"),
         units != "million pounds (GBP)") %>%
  ggplot() +
  geom_col(aes(x = reorder(area_visited, -value), y = value, fill = region_of_residence), position = "dodge") +
   theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 55, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Area Visited",
       y = "Thousands",
       fill = "",
       title = "Nights stayed and Visits made to the top Areas visted",
       subtitle = "2013 - 2019") +
  scale_fill_manual(values = c("England" = "#B20000", "Scotland" = "#211551"))+
  facet_wrap(units ~ .)
```

## create graph for total expenditure by region of residence and area visited
```{r}
regional_visits %>% 
  drop_na(area_visited) %>% 
  filter(region_of_residence != "All of GB",
         units != "million pounds (GBP)") %>% 
  group_by(area_visited, region_of_residence, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(area_visited, -sum), y = sum, fill = units), position = "dodge") +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1))
```


#Geospacial
## create a heat map for expeniture by are visited
```{r}
library(sf)

#read in spatial local authority data and simplify to 1km
la_zones <- st_read(here::here("raw_data/Local_Authority_Boundaries_-_Scotland/pub_las.shp")) %>%
  st_simplify(preserveTopology = FALSE, dTolerance = 1000)
```

```{r}
#use merge visitor data to shape file
visits_la_geo <- la_zones %>%
  merge(regional_visits, by.x = "code", by.y = "feature_code", all = TRUE)
```



```{r}
visits_la_geo %>% 
          drop_na(area_visited) %>% 
          filter(region_of_residence != "All of GB",
                 units == "million pounds (GBP)",
                 date_code == "2017-2019")%>% 
          ggplot() +
          geom_sf(aes(fill = value), colour = "black") +
          theme(axis.text.x = element_text(face = "bold", size = 10),
                axis.text.y = element_text(face = "bold", size = 10),
                title =element_text(size=12, face='bold'),
                axis.title=element_text(size=12),
                legend.position = "bottom",
                legend.text = element_text(size = "8")) +
           scale_fill_viridis_c(option="viridis") +
           labs(title = "Expenditure by Area visited",
               subtitle = "2017-2019",
               fill = "GBP in millions")
```
# Accomodation

```{r}
accomm_occupancy %>% 
  group_by(date_code, accommodation_type_and_occupancy, weekday_weekend, size_of_accommodation, location) %>% 
  summarise(sum = sum(value)) %>% 
  pivot_longer(cols = weekday_weekend:location, names_to = "category", values_to = "status") %>% 
  filter(status != "All")
```

```{r}
accomm_occupancy%>% 
  filter(weekday_weekend != "All",
         date_code == "2019") %>% 
  group_by(date_code, accommodation_type_and_occupancy, weekday_weekend, size_of_accommodation, location) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(accommodation_type_and_occupancy, -sum), y = sum, fill = weekday_weekend), position = "dodge") +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1))
```

```{r}
accomm_occupancy%>% 
  filter(size_of_accommodation != "All",
         date_code == "2019") %>% 
  group_by(accommodation_type_and_occupancy, weekday_weekend, size_of_accommodation, location,) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(accommodation_type_and_occupancy, sum), y = sum, fill = size_of_accommodation), position = "dodge", colour = "white") +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1))
```

```{r}
accomm_occupancy%>% 
  filter(location != "All",
         date_code == 2011) %>% 
  group_by(accommodation_type_and_occupancy, weekday_weekend, size_of_accommodation, location,) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(accommodation_type_and_occupancy, -sum), y = sum, fill = location), position = "dodge", colour = "white") +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1))
```




# Trends
## create visualisations to look at trends over years for day visits
```{r}
day_visit_activities %>% 
  group_by(date_code, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_line(aes(x = date_code, y = sum, colour = units)) +
  facet_grid(units ~ .)
```

```{r}
day_visit_activities %>% 
  filter( units == "million pounds (GBP)",
          tourism_activity == "All") %>% 
  group_by(date_code, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_line(aes(x = date_code, y = sum), colour = "#003300") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Year",
       y = "Millions",
       fill = "",
       title = "Domestic Tourist Trends by Expenditure",
       subtitle = "2013 - 2019") +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016, 2017, 2018, 2019))
```

```{r}
day_visit_activities %>% 
  filter( units == "Million Visits",
          tourism_activity == "All") %>% 
  group_by(date_code, units) %>% 
  summarise(sum = sum(value)) %>% 
  ggplot() +
  geom_line(aes(x = date_code, y = sum), colour = "#B27300") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1),
        legend.position = "top",
        axis.text.y = element_text(face = "bold", angle = 25, size = 10),
        title =element_text(size=12, face="bold"),
        axis.title=element_text(size= 12)) +
  labs(x = "Year",
       y = "Millions",
       fill = "",
       title = "Domestic Tourist Trends by Visits",
       subtitle = "2013 - 2019") +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016, 2017, 2018, 2019))
```


## create visualisations to look at trends over years for regional visits
```{r}
regional_visits %>% 
  drop_na(area_visited) %>% 
  group_by(date_code) %>% 
  filter(region_of_residence != "All of GB",
        units == "million pounds (GBP)") %>% 
  mutate(units = fct_reorder(units, desc(value))) %>% 
  ggplot() +
  geom_col(aes(x = date_code, y = value, fill = region_of_residence), position = "dodge") +
  theme(axis.text.x = element_text(face = "bold", angle = 45, hjust = 1, vjust = 1))
```



